---
title: "NYC flights - data wrangling"
subtitle: "due Thursday, September 15 at 4:20pm"
output: 
  html_document:
    toc: true
    toc_float: true
    number_section: false
    highlight: tango
    theme: "cosmo"
link-citations: yes
editor_options: 
  markdown: 
    wrap: sentence
---

```{r load-pkg, message = FALSE}
library(tidyverse)
library(nycflights13)
```

To demonstrate data wrangling we will use a dataset of characteristics of all flights departing from New York City (JFK, LGA, EWR) in 2013. We will use the `flights` data frame that comes from the `nycflights13` package. 

We first explore the data a bit. Examine the documentation as well.

```{r}
glimpse(flights)
```

# Exercises

Practice using `dplyr` verbs to perform each of the following tasks. 

# `select()`

1. Select only the departure delay and arrival delay columns, and store this into a new object called delays

```{r}

```

1. Select all variables in `flights`, excluding departure delays

```{r}

```

1. Select just the first three variables in `flights`

```{r}

```

# `arrange()`

4. Arrange the `flights` data by descending departure delays, with large departure delays on top.

```{r}

```

# `slice()`

5. Select (`slice`) the first 5 rows of `flights`

```{r}

```

6. Combine `arrange` and `slice` to select the 5 flights with the longest departure delays

```{r}

```


# `filter()`

7. Select all rows where the destination airport is "LAX"

```{r}

```

8. Select all rows where the destination airport is "LAX" and the arrival delay is less than 0.

```{r}

```

9. Select all rows where the destination airport is "LAX", "BUR", or "LGB" AND the departure or the arrival delay is negative. 

```{r}

```


# `mutate()`

10. `airtime` is provided in minutes. Create a new variable called `hours` that coverts `airtime` to hours. Make sure to save this new variable by overwriting the `flights` dataframe.

```{r}

```

11. Create a new variable called `mph` that computes the miles traveled (`distance`) per hour. Make sure to save this variable.

```{r}

```

12. Create a new variable called `on_time`  that is `TRUE` if the flight is on time and `FALSE` if it is not. Make sure to save this variable. 

```{r}

```

# `summarise()` and `group_by()`

13. Calculate the average departure delay

```{r}

```

14. Calculate the average departure delay for each month

```{r}

```

15. Count the number of on-time flights for each month

```{r}

```


16. Compute the proportion of flights that are on-time for each month

```{r}

```

17. In addition to the proportion on time for each month, also find the maximum delay, median delay, and the count of flights in each month.

```{r}

```

18. What is the proportion on time for flights traveling to "LAX", "BUR", and "LGB"?

```{r}

```

# More practice

19. Create a new dataset that only contains flights that do not have a missing departure time. Include the columns `year`,` month`, `day`, `dep_time`, `dep_delay`, and `dep_delay_hours` (the departure delay in hours). Note you may need to use `mutate()` to make one or more of these variables.

```{r}

```

20. For each airplane (uniquely identified by `tailnum`), use a `group_by()` paired with `summarize()` to find the sample size, mean, and standard deviation of flight distances. Then, pick off the top 5 and bottom 5 airplanes in terms of mean distance traveled per flight.

```{r}

```


# Submission

To submit, you need to export your .html document from RStudio Cloud and then upload it to the AE 03 assignment on Canvas. Refer back to [Lab 01](https://kgfitzgerald.github.io/APU_MATH_250/labs/lab-01-hello-r.html) for detailed instructions on how to export and submit your assignment. 

# Grading (10 pts)
 
Application exercises are graded on completion. You will receive full credit if you turn in the .html on time and a reasonable attempt was made on each exercise. 


